---
title: "Model Complexity for Gradient Boosting"
author: "McKade Thomas"
date: "7/11/2022"
output:
  pdf_document:
    extra_dependencies: ["float"]
---

```{r init, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.pos = "!H", fig.show="hold", out.extra = "")
```

```{r setup, include = FALSE}
source('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/ordinal_helper_methods.R')
source('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/main_helper_methods.R')
library(glmnet)
library(pROC)
library(MLmetrics)
library(corrplot)
library(kableExtra)
library(gbm)
library(fastDummies)
library(parallel)
library(tidyverse)
library(rmatio)
library(foreach)
library(doParallel)
library(ggplot2)
library(gridExtra)
library(corrplot)
library(randomForest)
library(arules)

# discretize(steel_encoded$Usage_kWh, breaks = 5)

## GLOBALS
MC_run = 50
MAX = 4
LEN = 40
```

# Steel Industry Energy Consumption Data (n=35040)
URL for data: https://archive.ics.uci.edu/ml/datasets/Steel+Industry+Energy+Consumption+Dataset
URL related work: https://www.tandfonline.com/doi/pdf/10.1080/09613218.2020.1809983?casa_token=VodoKalAs6AAAAAA:z6W4_oqbI5XKj41VsyKPs8FYZVnXyYoUBYhtgdTnFcKdj3V9BCXcYcgvkUbGadB-435p7MiF2w
# Goal for RMSE from paper: 5.69
```{r}
# Read in the data
steel <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/steel/steel.csv', sep = ",", header = TRUE)[-1]

# # steel$Usage_kWh <- log(steel$Usage_kWh)
# # steel <- steel[is.finite(steel$Usage_kWh),]
# # steel$Usage_kWh <- abs(steel$Usage_kWh - mean(steel$Usage_kWh))
# steel$rank <- order(steel$Usage_kWh)
# steel_binned <- bin_df(steel, "rank")
# steel_encoded <- steel_binned[,c(-(ncol(steel_binned)-1),-ncol(steel_binned))]
# cols_to_encode = c("WeekStatus", "Day_of_week", "Load_Type")
# steel_encoded <- dummy_cols(steel_encoded, select_columns = cols_to_encode) %>%
#                         dplyr::select(-all_of(cols_to_encode))
# cols_to_convert <- names(steel_encoded[,9:20])
# # cols_to_convert <- c("WeekStatus", "Day_of_week", "Load_Type")
# steel_encoded[cols_to_convert] <- lapply(steel_encoded[cols_to_convert], factor)
# steel_encoded <- steel_encoded[,c(-5,-19)]
# # cols_to_scale <- c("Lagging_Current_Reactive.Power_kVarh","Leading_Current_Reactive_Power_kVarh", "NSM",
# #                   "Lagging_Current_Power_Factor", "Leading_Current_Power_Factor")
# # steel_encoded[cols_to_scale] <- lapply(steel_encoded[cols_to_scale], scale)


steel <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/steel/steel.csv', sep = ",", header = TRUE)[-1]
# steel$rank <- rank(steel$Usage_kWh)
# steel_binned <- bin_df(steel, "rank")

steel_binned <- bin_df(steel, "Usage_kWh")
steel_encoded <- steel_binned[,c(-(ncol(steel_binned)-1),-ncol(steel_binned))]
steel_encoded$WeekStatus <- as.factor(steel_encoded$WeekStatus)
steel_encoded$Day_of_week <- as.factor(steel_encoded$Day_of_week)
steel_encoded$Load_Type <- as.factor(steel_encoded$Load_Type)

orchestrate_sim(dataset = steel_encoded, cont_response = "Usage_kWh", sample_size = 1000, MC = MC_run, 
                mc_min_trees = 0, mc_max_trees = MAX, full = steel_encoded, mc_tree_length = LEN) 
```


```{r}
# Read in the data
steel <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/steel/steel.csv', sep = ",", header = TRUE)[-1]


## RANKING
# steel$rank <- rank(steel$Usage_kWh)
steel_binned <- bin_df(steel, "Usage_kWh", num_bins = 10)
steel_encoded <- steel_binned[,c(-(ncol(steel_binned)-1),-ncol(steel_binned))]
steel_encoded$WeekStatus <- as.factor(steel_encoded$WeekStatus)
steel_encoded$Day_of_week <- as.factor(steel_encoded$Day_of_week)
steel_encoded$Load_Type <- as.factor(steel_encoded$Load_Type)

orchestrate_sim(dataset = steel_encoded, cont_response = "Usage_kWh", sample_size = 1000, MC = MC_run, 
                mc_min_trees = 0, mc_max_trees = MAX, full = steel_encoded, num_classes = 10, mc_tree_length = LEN)
```





\newpage
# Superconductivity Data

related work: https://arxiv.org/pdf/1803.10260.pdf
# Goal for RMSE from Paper: 9.4
```{r turbine}
conduct <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/conduct/train.csv', sep = ",", header = TRUE)
response <- "critical_temp"
conduct_names <- c(
  response, "range_ThermalConductivity","wtd_std_ThermalConductivity","range_atomic_radius",
  "wtd_gmean_ThermalConductivity","std_ThermalConductivity","wtd_entropy_Valence",
  "wtd_std_ElectronAffinity","wtd_entropy_atomic_mass","wtd_mean_Valence",
  "wtd_gmean_ElectronAffinity","wtd_range_ElectronAffinity","wtd_mean_ThermalConductivity",
  "wtd_gmean_Valence","std_atomic_mass","std_Density","wtd_entropy_ThermalConductivity",
  "wtd_range_ThermalConductivity","wtd_mean_atomic_mass","wtd_std_atomic_mass","gmean_Density"
)
conduct <- conduct[,(names(conduct) %in% conduct_names)]

# conduct$rank <- rank(conduct$critical_temp)
# conduct_binned <- bin_df(conduct, "rank")

conduct_binned <- bin_df(conduct, "critical_temp")
conduct_encoded <- conduct_binned[,c(-(ncol(conduct_binned)-1),-ncol(conduct_binned))]
# conduct_encoded <- conduct_encoded[,!((names(conduct_encoded) %in% c("wtd_mean_Valence", "wtd_entropy_Valence",
#                                                   "range_ThermalConductivity", "std_ThermalConductivity",
#                                                   "std_atomic_mass", "range_atomic_radius",
#                                                   "wtd_range_ThermalConductivity", "gmean_Density",
#                                                   "wtd_std_ThermalConductivity", "wtd_gmean_Valence",
#                                                   "wtd_range_ElectronAffinity")))]

orchestrate_sim(dataset = conduct_encoded, cont_response = "critical_temp", sample_size = 1000, MC = MC_run, 
                mc_min_trees = 0, mc_max_trees = 4, mc_tree_length = LEN, full = conduct_encoded)
```



```{r}
# conduct <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/conduct/train.csv', sep = ",", header = TRUE)
# response <- "critical_temp"
# conduct_names <- c(
#   response, "range_ThermalConductivity","wtd_std_ThermalConductivity","range_atomic_radius",
#   "wtd_gmean_ThermalConductivity","std_ThermalConductivity","wtd_entropy_Valence",
#   "wtd_std_ElectronAffinity","wtd_entropy_atomic_mass","wtd_mean_Valence",
#   "wtd_gmean_ElectronAffinity","wtd_range_ElectronAffinity","wtd_mean_ThermalConductivity",
#   "wtd_gmean_Valence","std_atomic_mass","std_Density","wtd_entropy_ThermalConductivity",
#   "wtd_range_ThermalConductivity","wtd_mean_atomic_mass","wtd_std_atomic_mass","gmean_Density"
# )
# conduct <- conduct[,(names(conduct) %in% conduct_names)]
# # conduct$critical_temp <- log(conduct$critical_temp)
# conduct_binned <- bin_df(conduct, response, num_bins = 10)
# conduct_encoded <- conduct_binned[,c(-(ncol(conduct_binned)-1),-ncol(conduct_binned))]

conduct <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/conduct/train.csv', sep = ",", header = TRUE)
response <- "critical_temp"
conduct_names <- c(
  response, "range_ThermalConductivity","wtd_std_ThermalConductivity","range_atomic_radius",
  "wtd_gmean_ThermalConductivity","std_ThermalConductivity","wtd_entropy_Valence",
  "wtd_std_ElectronAffinity","wtd_entropy_atomic_mass","wtd_mean_Valence",
  "wtd_gmean_ElectronAffinity","wtd_range_ElectronAffinity","wtd_mean_ThermalConductivity",
  "wtd_gmean_Valence","std_atomic_mass","std_Density","wtd_entropy_ThermalConductivity",
  "wtd_range_ThermalConductivity","wtd_mean_atomic_mass","wtd_std_atomic_mass","gmean_Density"
)
conduct <- conduct[,(names(conduct) %in% conduct_names)]
# conduct$rank <- rank(conduct$critical_temp)
# conduct_binned <- bin_df(conduct, "rank", num_bins = 10)
conduct_binned <- bin_df(conduct, "critical_temp", num_bins = 10)
conduct_encoded <- conduct_binned[,c(-(ncol(conduct_binned)-1),-ncol(conduct_binned))]
# conduct_encoded <- conduct_encoded[,!((names(conduct_encoded) %in% c("wtd_mean_Valence", "wtd_entropy_Valence",
#                                                   "range_ThermalConductivity", "std_ThermalConductivity",
#                                                   "std_atomic_mass", "range_atomic_radius",
#                                                   "wtd_range_ThermalConductivity", "gmean_Density",
#                                                   "wtd_std_ThermalConductivity", "wtd_gmean_Valence",
#                                                   "wtd_range_ElectronAffinity")))]


orchestrate_sim(dataset = conduct_encoded, cont_response = "critical_temp", sample_size = 1000, MC = 20, 
                mc_min_trees = 0, mc_max_trees = 3.5, mc_tree_length = LEN, full = conduct_encoded, num_classes = 10)
```






# Concrete
URL: https://data.world/uci/concrete-compressive-strength
Related Work: https://www.researchgate.net/publication/222447231_Modeling_of_Strength_of_High-Performance_Concrete_Using_Artificial_Neural_Networks_Cement_and_Concrete_research_2812_1797-1808
# Goal for R2 from Paper: 0.91
```{r}
concrete <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/concrete/concrete.csv', sep = ",", header = TRUE)
names <- c("cement", "blast", "fly_ash", "water", "plastic", "coarse", "fine", "age", "strength")
colnames(concrete) <- names

# concrete$rank <- rank(concrete$strength)
# concrete_binned <- bin_df(concrete, "rank")

concrete_binned <- bin_df(concrete, "strength")
concrete_encoded <- concrete_binned[,c(-(ncol(concrete_binned)-1),-ncol(concrete_binned))]
# concrete_encoded <- concrete_encoded[,!(names(concrete_encoded) %in% "rank")]

orchestrate_sim(dataset = concrete_encoded, cont_response = "strength", sample_size = 600, MC = 50, 
                mc_min_trees = 0, mc_max_trees = MAX, full = concrete_encoded, mc_tree_length = LEN)
```


# Concrete 10 classes
```{r}
concrete <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/concrete/concrete.csv', sep = ",", header = TRUE)
names <- c("cement", "blast", "fly_ash", "water", "plastic", "coarse", "fine", "age", "strength")
colnames(concrete) <- names

# concrete$rank <- rank(concrete$strength)
concrete_binned <- bin_df(concrete, "strength", num_bins = 10)
concrete_encoded <- concrete_binned[,c(-(ncol(concrete_binned)-1),-ncol(concrete_binned))]
# concrete_encoded <- concrete_encoded[,!(names(concrete_encoded) %in% "rank")]

orchestrate_sim(dataset = concrete_encoded, cont_response = "strength", sample_size = 600, MC = 50, 
                mc_min_trees = 0, mc_max_trees = MAX, full = concrete_encoded, num_classes = 10, mc_tree_length = LEN)
```





# Evaluate Model Performance
```{r}
cont_response = "Usage_kWh"
steel <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/steel/steel.csv', sep = ",", header = TRUE)[-1]

# steel$rank <- order(steel$Usage_kWh)
steel_binned <- bin_df(steel, "Usage_kWh")
steel_encoded <- steel_binned[,c(-(ncol(steel_binned)-1),-ncol(steel_binned))]
steel_encoded$WeekStatus <- as.factor(steel_encoded$WeekStatus)
steel_encoded$Day_of_week <- as.factor(steel_encoded$Day_of_week)
steel_encoded$Load_Type <- as.factor(steel_encoded$Load_Type)
# steel_encoded <- steel_encoded[,!(names(steel_encoded) %in% "rank")]

# steel_binned <- bin_df(steel, "Usage_kWh")
# steel_encoded <- steel_binned[,c(-(ncol(steel_binned)-1),-ncol(steel_binned))]
# cols_to_encode = c("WeekStatus", "Day_of_week", "Load_Type")
# steel_encoded <- dummy_cols(steel_encoded, select_columns = cols_to_encode) %>%
#                         dplyr::select(-all_of(cols_to_encode))
# cols_to_convert <- names(steel_encoded[,9:20])
# # cols_to_convert <- c("WeekStatus", "Day_of_week", "Load_Type")
# steel_encoded[cols_to_convert] <- lapply(steel_encoded[cols_to_convert], factor)
# steel_encoded <- steel_encoded[,c(-5,-19)]

# ind <- sample(1:nrow(steel_encoded), 1000)
sub <- sample_by_distribution_ten(steel_encoded)
train <- sub[[1]]
test <- sub[[2]]
# test_ranks = test$rank
# train <- train[,names(train) != "rank"]
# test <- test[,names(test) != "rank"]

mdl <- gbm(as.numeric(class) ~ ., data = train[, names(train) != cont_response], n.trees = 100)

test_preds <- predict(mdl, newdata = test[, !(names(test) %in% c("class",cont_response))],
                          n.trees = 100)

cont_test_preds <- transform_ranks(test_preds, test, nrows = nrow(steel_encoded))

sqrt(MSE(y_pred = cont_test_preds, y_true = test_ranks))
test$preds <- cont_test_preds
sqrt(mean((test_ranks-mean(test_ranks))^2))


ggplot(test, aes(x = test_ranks, y=preds, col = class)) + geom_point() +
  labs(x="Actual",
       y="Predicted") +
  theme_classic()
```






# Exploratory Data Analysis
```{r}
steps = c(1,2,3,4,5)
ggplot(as.data.frame(steps), aes(x = seq_along(steps), y = steps)) + 
  geom_step(col = "grey") + geom_point(col = "grey") +
  geom_smooth(aes(x=steps, y=steps-.5)) +
  theme_classic() +
  labs(x = "x",
       y = "y")
```



```{r}
# Class Distribution
p1 <- ggplot(steel_encoded, aes(x=class, y=Usage_kWh, fill = class)) + 
  geom_boxplot() + 
  theme_classic()+ 
  theme(legend.position = "none") +
  labs(title = "Steel",
       x = "Class",
       y = "Energy Usage")
  
p2 <- ggplot(conduct_encoded, aes(x=class, y=critical_temp, fill = class)) +
  geom_boxplot() + 
  theme_classic()+ 
  theme(legend.position = "none") +
  labs(title = "Superconductivity",
       x = "Class",
       y = "Critical Temp")
p3 <- ggplot(concrete_encoded, aes(x=class, y=strength, fill = class)) + 
  geom_boxplot() + 
  theme_classic() + 
  theme(legend.position = "none") +
  labs(title = "Concrete",
       x = "Class",
       y = "Compression Strength")

# 
# ggplot(steel_encoded, aes(x=class)) +
#   geom_boxplot(aes(y=Usage_kWh)) +
#   geom_boxplot(conduct_encoded, aes(y=critical_temp))
# 
# 
# comb_boxplot <- as.data.frame(
#   cbind(steel_encoded$class, 
#         steel_encoded$Usage_kWh, 
#         conduct_encoded$critical_temp,
#         concrete_encoded$strength)
# )
# colnames(comb_boxplot) <- c("Class", "Usage", "Critical_Temp", "Strength")

grid.arrange(p1, p2, p3, nrow=1)
```



```{r}
library(reshape)
steel_melt <- melt(steel[2:7])
conduct_melt <- melt(conduct_encoded[,!(names(conduct_encoded) %in% c("class", "critical_temp"))])
concrete_melt <- melt(concrete[1:8])

plot_features(steel_melt, title = "Steel")
plot_features(conduct_melt, title = "Conduct")
plot_features(concrete_melt, title = "Concrete")

plot_features(conduct_melt, title = "Conduct")
```


```{r}
ggplot(steel_encoded, aes(x = Usage_kWh)) + 
  geom_histogram(aes(y =..density..), colour = "black", fill = "grey", bins = 65) + 
  geom_density(alpha=.2, fill="red", color = "grey") +
  theme_classic() + 
  labs(title = "Distribution of Energy Usage (kWh)",
       y = "Density", x = "Energy Usage")

ggplot(steel_encoded, aes(x = Usage_kWh)) + 
  geom_density(alpha=.2, fill="red", color = "grey") +
  geom_density(aes(x = log(Usage_kWh), y = stat(density)),alpha=.2, fill="blue", color = "grey") +
  # ylim(0, 0) +
  xlim(0, 10) +
  theme_classic() + 
  labs(title = "Logged Vs. Unlogged Energy Usage (kWh)",
       y = "Density", x = "Compression Strength") +
  theme(axis.title.y.right = element_text(color = "blue"))



ggplot(conduct_encoded, aes(x = critical_temp)) + 
  geom_histogram(aes(y =..density..), colour = "black", fill = "grey", bins = 65) + 
  geom_density(alpha=.2, fill="red", color = "grey") +
  theme_classic() + 
  labs(title = "Distribution of Critical Temperature",
       y = "Density", x = "Critical Temp")

ggplot(conduct_encoded, aes(x = critical_temp)) + 
  geom_density(alpha=.2, fill="red", color = "grey") +
  geom_density(aes(x = log(critical_temp), y = stat(density)),alpha=.2, fill="blue", color = "grey") +
  # ylim(0, 0) +
  xlim(-5, 35) +
  theme_classic() + 
  labs(title = "Logged Vs. Unlogged Critical Temp",
       y = "Density", x = "Compression Strength") +
  theme(axis.title.y.right = element_text(color = "blue"))



ggplot(gas_encoded, aes(x = strength)) + 
  geom_histogram(aes(y =stat(density)), colour = "black", fill = "grey", bins = 50) + 
  geom_density(alpha=.2, fill="red", color = "grey") +
  theme_classic() + 
  labs(title = "Distribution of Compression Strength",
       y = "Density", x = "Compression Strength")

ggplot(gas_encoded, aes(x = strength)) + 
  geom_density(alpha=.2, fill="red", color = "grey") +
  geom_density(aes(x = log(strength), y = stat(density)),alpha=.2, fill="blue", color = "grey") +
  xlim(-5, 50) +
  theme_classic() + 
  labs(title = "Logged Vs. Unlogged Compression Strength",
       y = "Density", x = "Compression Strength") +
  theme(axis.title.y.right = element_text(color = "blue"))
```




```{r, correlation plots}
plot_steel <- steel
colnames(plot_steel) <- c("Usage", "Lag_Reactive", "Lead_Reactive", "tCO2", 
                          "Lag_Factor", "Lead_Factor", "NSM", "WeekStatus",
                          "Weekday", "Load_Type")
plot_steel <- plot_steel[,!(names(plot_steel) %in% c("Lead_Reactive"))]
corrplot(cor(plot_steel[2:6]), method="shade", tl.col = "grey", tl.cex = 0.8, tl.srt = 40)
cor(plot_steel[2:6]) %>% kbl(caption = "") %>% 
           kable_classic(full_width = F, html_font = "Cambria", latex_options = "hold_position")

plot_conduct <- conduct[,!((names(conduct) %in% c("wtd_mean_Valence", "wtd_entropy_Valence",
                                                  "range_ThermalConductivity", "std_ThermalConductivity",
                                                  "std_atomic_mass", "range_atomic_radius",
                                                  "wtd_range_ThermalConductivity", "gmean_Density",
                                                  "wtd_std_ThermalConductivity", "wtd_gmean_Valence",
                                                  "wtd_range_ElectronAffinity", "critical_temp")))]

corrplot(cor(plot_conduct), method="shade", tl.col = "grey", tl.cex = 0.6, tl.srt = 50)
cor(plot_conduct) %>% kbl(caption = "") %>% 
           kable_classic(full_width = F, html_font = "Cambria", latex_options = "hold_position")



corrplot(cor(concrete[1:8]), method="shade", tl.col = "grey")
cor(concrete[1:8]) %>% kbl(caption = "") %>% 
           kable_classic(full_width = F, html_font = "Cambria", latex_options = "hold_position")
```


##TODO: Fix model equation on line 564
##TODO: Include RF as baseline naive model for comparing performance improvement
##TODO: Fix boxplots/tables for input features for superconductivity
##TODO: Superconductivity feature selection
##TODO: Adjust RF to transform predict continuous
##TODO: Calculate percent MSE change based on RF


\newpage

# Appendix: All code for this report

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```
