---
title: "deprecated_thesis_code"
author: "McKade Thomas"
date: "10/11/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r crime_names, include = FALSE}
true_names <- c(
 "CrimeRate"
 ,"state" 
 ,"county" 
 ,"community" 
 , "communityname" 
 ,"fold" 
 ,"population" 
 ,"householdsize" 
 ,"racepctblack" 
 ,"racePctWhite" 
 ,"racePctAsian" 
 ,"racePctHisp" 
 ,"agePct12t21" 
 ,"agePct12t29" 
 ,"agePct16t24" 
 ,"agePct65up" 
 ,"numbUrban" 
 ,"pctUrban" 
 ,"medIncome" 
 ,"pctWWage" 
 ,"pctWFarmSelf" 
 ,"pctWInvInc" 
 ,"pctWSocSec" 
 ,"pctWPubAsst" 
 ,"pctWRetire" 
 ,"medFamInc" 
 ,"perCapInc" 
 ,"whitePerCap" 
 ,"blackPerCap" 
 ,"indianPerCap" 
 ,"AsianPerCap" 
 ,"OtherPerCap" 
 ,"HispPerCap" 
 ,"NumUnderPov" 
 ,"PctPopUnderPov" 
 ,"PctLess9thGrade" 
 ,"PctNotHSGrad" 
 ,"PctBSorMore" 
 ,"PctUnemployed" 
 ,"PctEmploy" 
 ,"PctEmplManu" 
 ,"PctEmplProfServ" 
 ,"PctOccupManu" 
 ,"PctOccupMgmtProf" 
 ,"MalePctDivorce" 
 ,"MalePctNevMarr" 
 ,"FemalePctDiv" 
 ,"TotalPctDiv" 
 ,"PersPerFam" 
 ,"PctFam2Par" 
 ,"PctKids2Par" 
 ,"PctYoungKids2Par" 
 ,"PctTeen2Par" 
 ,"PctWorkMomYoungKids" 
 ,"PctWorkMom" 
 ,"NumIlleg" 
 ,"PctIlleg" 
 ,"NumImmig" 
 ,"PctImmigRecent" 
 ,"PctImmigRec5" 
 ,"PctImmigRec8" 
 ,"PctImmigRec10" 
 ,"PctRecentImmig" 
 ,"PctRecImmig5" 
 ,"PctRecImmig8" 
 ,"PctRecImmig10" 
 ,"PctSpeakEnglOnly" 
 ,"PctNotSpeakEnglWell" 
 ,"PctLargHouseFam" 
 ,"PctLargHouseOccup" 
 ,"PersPerOccupHous" 
 ,"PersPerOwnOccHous" 
 ,"PersPerRentOccHous" 
 ,"PctPersOwnOccup" 
 ,"PctPersDenseHous" 
 ,"PctHousLess3BR" 
 ,"MedNumBR" 
 ,"HousVacant" 
 ,"PctHousOccup" 
 ,"PctHousOwnOcc" 
 ,"PctVacantBoarded" 
 ,"PctVacMore6Mos" 
 ,"MedYrHousBuilt" 
 ,"PctHousNoPhone" 
 ,"PctWOFullPlumb" 
 ,"OwnOccLowQuart" 
 ,"OwnOccMedVal" 
 ,"OwnOccHiQuart" 
 ,"RentLowQ" 
 ,"RentMedian" 
 ,"RentHighQ" 
 ,"MedRent" 
 ,"MedRentPctHousInc" 
 ,"MedOwnCostPctInc" 
 ,"MedOwnCostPctIncNoMtg" 
 ,"NumInShelters" 
 ,"NumStreet" 
 ,"PctForeignBorn" 
 ,"PctBornSameState" 
 ,"PctSameHouse85" 
 ,"PctSameCity85" 
 ,"PctSameState85" 
 ,"LemasSwornFT" 
 ,"LemasSwFTPerPop" 
 ,"LemasSwFTFieldOps" 
 ,"LemasSwFTFieldPerPop" 
 ,"LemasTotalReq" 
 ,"LemasTotReqPerPop" 
 ,"PolicReqPerOffic" 
 ,"PolicPerPop" 
 ,"RacialMatchCommPol" 
 ,"PctPolicWhite" 
 ,"PctPolicBlack" 
 ,"PctPolicHisp" 
 ,"PctPolicAsian" 
 ,"PctPolicMinor" 
 ,"OfficAssgnDrugUnits" 
 ,"NumKindsDrugsSeiz" 
 ,"PolicAveOTWorked" 
 ,"LandArea" 
 ,"PopDens" 
 ,"PctUsePubTrans" 
 ,"PolicCars" 
 ,"PolicOperBudg" 
 ,"LemasPctPolicOnPatr" 
 ,"LemasGangUnitDeploy" 
 ,"LemasPctOfficDrugUn" 
 ,"PolicBudgPerPop")
```



# Bike Sharing Data (n=731)
URL for data: https://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset
```{r bike, echo = FALSE, message = FALSE}
# Read in the data
bike <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/bike/day.csv', sep = ",", header = T)[c(-1,-2)]

bike_binned <- bin_df(bike, "cnt")
bike_encoded <- bike_binned[,c(-(ncol(bike_binned)-1),-ncol(bike_binned))]
train_index <- sample(1:nrow(bike_encoded), nrow(bike_encoded)*0.7)
bike_train <- bike_encoded[train_index,]
bike_test <- bike_encoded[-train_index,]
return_df <- run_sim(train = bike_train, test = bike_test, cont_response = "cnt", cont_location = 15)


plot_return(return_df)
```


# Bike Sharing Data WITH NOISE
```{r bike_noise, echo = FALSE, message = FALSE}
for(noise in NOISES){
  return_df <- run_sim(train = add_noise(bike_train, "cnt", noise),
          test = add_noise(bike_test, "cnt", noise),
        cont_response = "cnt", cont_location = 15, 
        noise_added = TRUE, noise_factor = noise)
  plot_return(return_df)
}
```


\newpage

# Music Data (n=3943)
```{r music, echo = FALSE, message = FALSE}
music <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/music/music.txt', sep = ",", header = F)
music_binned <- bin_df(music, "V61")
music_encoded <- music_binned[,c(-(ncol(music_binned)-1),-ncol(music_binned))]
train_index <- sample(1:nrow(music_encoded), nrow(music_encoded)*0.7)
music_train <- music_encoded[train_index,]
music_test <- music_encoded[-train_index,]
return_df <- run_sim(train = music_train, test = music_test, cont_response = "V61", cont_location = 62)
plot_return(return_df)
```

# Music Data WITH NOISE
```{r music_noise, echo = FALSE, message = FALSE}
for(noise in NOISES){
  return_df <- run_sim(train = add_noise(music_train, variable = "V61", noise),
          test = add_noise(music_test, variable = "V61", noise),
        cont_response = "V61", cont_location = 62, 
        noise_added = TRUE, noise_factor = noise)
  plot_return(return_df)
}
```


# Abalone Data (n=4176)
URL for data: https://archive.ics.uci.edu/ml/datasets/abalone
Related Work: https://www.researchgate.net/profile/Willem-Waegeman-2/publication/228350235_An_ensemble_of_weighted_support_vector_machines_for_ordinal_regression/links/0046353be578c3aa0b000000/An-ensemble-of-weighted-support-vector-machines-for-ordinal-regression.pdf
```{r abalone, echo = FALSE, message = FALSE}
# Read in the data
aba <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/abalone/abalone.data', sep = ",", header = T)
true_aba_names <- c("Sex", "Length", "Diameter", "Height", "Whole_wt", 
                    "Shucked_wt", "Viscera_wt", "Shell_wt", "Rings")
colnames(aba) <- true_aba_names

aba_binned <- bin_df(aba, "Rings")
aba_features <- aba_binned[,c(-11,-12)]
aba_encoded <- dummy_cols(aba_features, select_columns = 'Sex') %>% dplyr::select(-Sex)

MC <- 100
mc_tree_length = 30
mc_min_trees = 1
mc_max_trees = 4
mc_tree_set <- unique(round(10^(seq(from=mc_min_trees, to=mc_max_trees, length=mc_tree_length))))
binned_best_mat <- matrix(NA, nrow = MC, ncol = length(mc_tree_set))
cont_best_mat <- matrix(NA, nrow = MC, ncol = length(mc_tree_set))

full_start_time = Sys.time()
for (i in 1:MC) {
  start_time = Sys.time()
  cat(glue('Begin Monte Carlo Iteration {i} of {MC}'), '\n')
  sub_sample <- sample_by_distribution(aba_encoded, size = 1000)
  return_df <- run_sim(train = sub_sample[[1]],
                       test = sub_sample[[2]],
                       cont_response = "Rings",
                       min_trees = mc_min_trees,
                       max_trees = mc_max_trees,
                       tree_length = mc_tree_length,
                       plot = FALSE)

  binned_best_mat[i,] <- return_df$test_mat
  cont_best_mat[i,] <- return_df$test_cont_mat

  cat(glue('--End Iteration {i} of {MC}: Completed in {round(Sys.time() - start_time,2)}'),'\n')
}
cat(glue('Full Run Success: Completed in {round(Sys.time() - full_start_time,2)}','\n'))
plot_return_mc(mc_tree_set, binned_best_mat, cont_best_mat)
```


\newpage

# Abalone Data WITH NOISE
```{r abalone_noise, echo = FALSE, message = FALSE}
for(noise in NOISES){
  return_df <- run_sim(train = add_noise(aba_train, variable = "Rings", noise), 
          test = add_noise(aba_test, variable = "Rings", noise),
        cont_response = "Rings", cont_location = 10, 
        noise_added = TRUE, noise_factor = noise)
  plot_return(return_df)
}
```


# Appliances Energy Data (n=19735)
URL for dtaa: https://archive.ics.uci.edu/ml/datasets/Appliances+energy+prediction
```{r energy, echo = FALSE, message = FALSE}
# Read in the data
energy <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/energy/energy.csv', sep = ",", header = T)[-1]

energy_binned <- bin_df(energy, "Appliances")
energy_encoded <- energy_binned[,c(-30,-31)]
energy_sizes = c(100,1000,7000)
for(size in energy_sizes){
  return_df <- run_sim(train = sample_by_distribution(energy_encoded, size = size),
          test = energy_encoded,
        cont_response = "Appliances", cont_location = 2)
  plot_return(return_df)
}
```


\newpage



# Appliances Energy Data WITH NOISE
```{r energy_noise, echo = FALSE, message = FALSE}
energy_sampled <- sample_by_distribution(energy_encoded, size = 1500)
for(noise in NOISES){
  return_df <- run_sim(train = add_noise(energy_sampled, "Appliances", noise),
          test = add_noise(energy_encoded, "Appliances", noise),
        cont_response = "Appliances", cont_location = 2, 
        noise_added = TRUE, noise_factor = noise)
  plot_return(return_df)
}

```


\newpage

\newpage

# Census Data WITH NOISE
```{r census_noise, echo = FALSE, message = FALSE}
census_sampled <- sample_by_distribution(census_encoded, size = 700)
for(noise in NOISES){
  run_sim(train = add_noise(census_sampled, "fnlgwt", noise), 
          test = add_noise(census_encoded, "fnlgwt", noise),
        cont_response = "fnlgwt", cont_location = 4, 
        noise_added = TRUE, noise_factor = noise)
}
```




```{r}
# library(e1071)
trees <- 500
# steel_encoded <- steel_encoded[is.finite(steel_encoded$Usage_kWh),]
steel_sub_sample <- sample_by_distribution(steel_encoded[,c(-5,-19)], size = 26000)
train <- steel_sub_sample[[1]]
test <- steel_sub_sample[[2]]
MDL <- gbm(as.numeric(class) ~ ., data = train[, names(train) != "Usage_kWh"], 
             distribution = "gaussian", n.trees = trees, n.cores = 8, 
             shrinkage = 0.1, interaction.depth = 1, keep.data = FALSE)

train_preds <- predict(MDL, newdata = train[, (names(train) != "Usage_kWh") & 
                                              (names(train) != "class")],
                       n.trees = trees)

test_preds <- predict(MDL, newdata = test[, (names(test) != "Usage_kWh") & 
                                              (names(test) != "class")],
                       n.trees = trees)

# plot(x = as.numeric(train[,"class"]), 
#      y = train_preds,
#      main = glue('Train vs Pred for {trees} Trees'),
#      xlab = glue("True"),
#      ylab = glue("Train Predicted"))
# abline(coef=c(0,1))

plot(x = as.numeric(test[,"class"]), 
     y = test_preds,
     main = glue('Test vs Pred for {trees} Trees'),
     xlab = glue("True"),
     ylab = glue("Test Predicted"))
abline(coef=c(0,1))
MSE(y_pred = train_preds, y_true = as.numeric(train$class))
MSE(y_pred = test_preds, y_true = as.numeric(test$class))


binned_train_resids <- as.numeric(train$class) - train_preds
binned_test_resids <- as.numeric(test$class) - test_preds

cont_train_lm <- lm(train[,"Usage_kWh"] ~ binned_train_resids + train$class, data = train)
cont_test_lm <- lm(test[,"Usage_kWh"] ~ binned_test_resids + test$class, data = test)

train_cont_preds <- predict(cont_train_lm, type = "response")
test_cont_preds <- predict(cont_test_lm, type = "response")

# plot(x = train[,"Usage_kWh"], 
#      y = train_cont_preds,
#      main = glue('Train vs Pred for {trees} Trees'),
#      xlab = glue("True"),
#      ylab = glue("Predicted"))
# abline(coef=c(0,1))

plot(x = test[,"Usage_kWh"], 
     y = test_cont_preds,
     main = glue('Test vs Pred for {trees} Trees'),
     xlab = glue("True"),
     ylab = glue("Predicted"))
abline(coef=c(0,1))


sqrt(MSE(y_pred = train_cont_preds, y_true = train[,"Usage_kWh"]))
sqrt(MSE(y_pred = test_cont_preds, y_true = test[,"Usage_kWh"]))

```




```{r}
# aggregate(Usage_kWh~class, FUN=max, data=steel_encoded)
hist(steel_encoded$Usage_kWh, main = "Energy Usage (kWh)", 
     xlab = "Usage", breaks = 105)
abline(v=c(3.1,3.74,12.56,57.74,157.18), col='blue', lwd=0.5)

hist(1/steel_encoded$Usage_kWh, main = "Inverted Energy Usage (kWh)", 
     xlab = "Usage", breaks = 120)
abline(v=c(0, 0.32,0.26,0.08,0.02), col='blue', lwd=0.5)

hist(log(steel_encoded$Usage_kWh), main = "Logged Energy Usage (kWh)", 
     xlab = "Usage", breaks = 70)
abline(v=c(1.13,1.32,2.53,4.06,5.06), col='blue', lwd=0.5)



hist(gas_encoded$strength, main = "Binning Compression Strength", 
     xlab = "Usage", breaks = 55)
abline(v=c(2.33, 20.97, 30.96, 39.05, 50.53, 82.6), col='blue', lwd=0.5)



ggplot(gas_encoded, aes(x=strength)) + 
  geom_histogram(color = "black", fill = "grey", bins = 50) +
  theme_classic() +
  geom_vline(xintercept = 0.7, col = "blue", linetype = "dashed", size = 1.1) +
  geom_vline(xintercept = 20.97, col = "blue", linetype = "dashed", size = 1.1) +
  geom_vline(xintercept = 30.96, col = "blue", linetype = "dashed", size = 1.1) +
  geom_vline(xintercept = 39.05, col = "blue", linetype = "dashed", size = 1.1) +
  geom_vline(xintercept = 50.53, col = "blue", linetype = "dashed", size = 1.1) +
  geom_vline(xintercept = 84, col = "blue", linetype = "dashed", size = 1.1) +
  labs(title = "Binning Continuous Response",
       x = "Compression Strength",
       y = "Count")
  
```






# Temperature
```{r}
temp <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/temp/temperature.csv', sep = ",", header = TRUE)

temp <- drop_na(temp)
temp_binned <- bin_df(temp, "Next_Tmin")
temp_encoded <- temp_binned[,c(-(ncol(temp_binned)-1),-ncol(temp_binned))]
temp_encoded <- temp_encoded[,c(-3,-25)]


MC <- 5
mc_tree_length = 20
mc_min_trees = 1
mc_max_trees = 4.7
mc_tree_set <- unique(round(10^(seq(from=mc_min_trees, to=mc_max_trees, length=mc_tree_length))))
binned_best_mat <- matrix(NA, nrow = MC, ncol = length(mc_tree_set))
cont_best_mat <- matrix(NA, nrow = MC, ncol = length(mc_tree_set))

full_start_time = Sys.time()
for (i in 1:MC) {
  start_time = Sys.time()
  cat(glue('Begin Monte Carlo Iteration {i} of {MC}'), '\n')
  sub_sample <- sample_by_distribution(temp_encoded, size = 700)
  return_df <- run_sim(train = sub_sample[[1]],
                       test = sub_sample[[2]],
                       cont_response = "Next_Tmin",
                       min_trees = mc_min_trees,
                       max_trees = mc_max_trees,
                       tree_length = mc_tree_length,
                       plot = FALSE)

  binned_best_mat[i,] <- return_df$test_mat
  cont_best_mat[i,] <- return_df$test_cont_mat

  cat(glue('--End Iteration {i} of {MC}: Completed in {round(Sys.time() - start_time,2)}'),'\n')
}
cat(glue('Full Run Success: Completed in {round(Sys.time() - full_start_time,2)}','\n'))
plot_return_mc(mc_tree_set, binned_best_mat, cont_best_mat)
```












Turbine Data: https://archive.ics.uci.edu/ml/datasets/Gas+Turbine+CO+and+NOx+Emission+Data+Set
Related works: https://journals.tubitak.gov.tr/cgi/viewcontent.cgi?article=1505&context=elektrik
```{r}
# Read in the data
# t1 <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/turbine/gt_2011.csv', sep = ",", header = T)
# t2 <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/turbine/gt_2012.csv', sep = ",", header = T)
# t3 <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/turbine/gt_2013.csv', sep = ",", header = T)
# t4 <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/turbine/gt_2014.csv', sep = ",", header = T)
# t5 <- read.table('C:/Users/mckad/Documents/USU Coursework/Thesis Work/Ordinal Classification/datasets/turbine/gt_2015.csv', sep = ",", header = T)
# 
# turbine <- rbind(t1,t2,t3,t4,t5)
# # turbine$TEY <- log(turbine$TEY)
# 
# turbine_binned <- bin_df(turbine, "TEY")
# turbine_encoded <- turbine_binned[,c(-13,-14)]
# 
# SIZES = unique(round(10^(seq(from=1, to=3.5, length=20))))*5
# 
# MC <- 25
# mc_min_trees = 1
# mc_max_trees = 2.7
# binned_best_trees_mat <- matrix(NA, nrow = MC, ncol = length(SIZES))
# cont_best_trees_mat <- matrix(NA, nrow = MC, ncol = length(SIZES))
# 
# full_start_time = Sys.time()
# for(i in 1:MC){
#   start_time = Sys.time()
#   cat(glue('Begin Monte Carlo Iteration {i} of {MC}'), '\n')
#   for (j in 1:length(SIZES)){
#     cat(glue('----Begin Iteration {j} of {length(SIZES)} with n = {SIZES[j]}', '\n', '\n'))
#     size_sample <- sample_by_distribution(turbine_encoded, size = SIZES[j])
#     return_df <- run_sim(train = size_sample[[1]],
#                          test = size_sample[[2]],
#                          cont_response = "TEY",
#                          min_trees = mc_min_trees,
#                          max_trees = mc_max_trees,
#                          tree_length = 20,
#                          plot = FALSE)
# 
#     binned_thresh <- min(return_df$test_mat) * 1.01
#     cont_thresh <- min(return_df$test_cont_mat) * 1.01
# 
#     binned_best_trees_mat[i,j] <- min(return_df[which(return_df$test_mat <= binned_thresh),
#                                                 "tree_set"])
#     cont_best_trees_mat[i,j] <- min(return_df[which(return_df$test_cont_mat <= cont_thresh),
#                                                 "tree_set"])
#   }
#   cat(glue('End Iteration {i} of {MC}: Completed in {round(Sys.time() - start_time,2)}'),'\n')
# }
# cat(glue('Full Run Success: Completed in {round(Sys.time() - full_start_time,2)}','\n'))
# plot_best_tree_mc(return_df, binned_best_trees_mat, cont_best_trees_mat)
```


# Sources
Jitter function: https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/jitter